---
title: "Load Data"
author: "Corinne Strawser"
date: "2026-02-06"
output: html_document
---

Patient       Sample 	                Section   Timepoint   Data
Patient 18    NP-19-0036945_B1        B1 	      D0          /igm/projects/240320_GSL-KM-3829_KC_visium/count/B1/outs
Patient 18    NP-20-0005714_F1 	      F1 	      D122        /igm/projects/240320_GSL-KM-3829_KC_visium/count/F1/outs
Patient 22    NP-20-0025854_C1_Pre    C1        D0          /igm/projects/241014_GSL-LPRV-4138_Visium/counts/NP-20-0025854_C1_Pre/NP-20-0025854_C1_Pre/outs
Patient 22    NP-20-003393_A1_Post    A1        D51         /igm/projects/250912_GSL-LPRV-4558_C134_M032/counts/NP-20-003393_A1/outs
Patient 22    NP-20-0043174_D4_Post   D4        D119        /igm/projects/241014_GSL-LPRV-4138_Visium/counts/NP-20-0043174_D4_Post/NP-20-0043174_D4_Post/outs

# load data from each sample and combine into a list
```{r load_data}
pt18_pre <-
  Load10X_Spatial(data.dir = "/igm/projects/240320_GSL-KM-3829_KC_visium/count/B1/outs/",
                  filename = "filtered_feature_bc_matrix.h5",
                  slice = "oHSV_d0")
pt18_post <-
  Load10X_Spatial(data.dir = "/igm/projects/240320_GSL-KM-3829_KC_visium/count/F1/outs/",
                  filename = "filtered_feature_bc_matrix.h5",
                  slice = "oHSV_d122")
pt22_pre <-
  Load10X_Spatial(data.dir = "/igm/projects/241014_GSL-LPRV-4138_Visium/counts/NP-20-0025854_C1_Pre/NP-20-0025854_C1_Pre/outs",
                  filename = "filtered_feature_bc_matrix.h5",
                  slice = "oHSV_d0")
pt22_post1 <-
  Load10X_Spatial(data.dir = "/igm/projects/250912_GSL-LPRV-4558_C134_M032/counts/NP-20-003393_A1/outs",
                  filename = "filtered_feature_bc_matrix.h5",
                  slice = "oHSV_d51")
pt22_post2 <-
  Load10X_Spatial(data.dir = "/igm/projects/241014_GSL-LPRV-4138_Visium/counts/NP-20-0043174_D4_Post/NP-20-0043174_D4_Post/outs",
                  filename = "filtered_feature_bc_matrix.h5",
                  slice = "oHSV_d119")

objs <-
  list("pt18_pre" = pt18_pre,
       "pt18_post" = pt18_post,
       "pt22_pre" = pt22_pre,
       "pt22_post1" = pt22_post1,
       "pt22_post2" = pt22_post2)
```

# find percentage of mitochondrial genes and subset
```{r subset_data}
for (i in names(objs)){
  objs[[i]]$orig.ident <- i
  objs[[i]] <-
    objs[[i]] %>%
    PercentageFeatureSet(pattern = "^MT-",
                         assay = "Spatial",
                         col.name = "percent.mt") %>%
    subset(nCount_Spatial > 0)
  saveRDS(objs[[1]],
          paste0(rdata, i, ".rds"))
}
```

# analyze patient 18 and patient 22 independently, merging pre- and post-oHSV samples from each patient
```{r}
# make a list of the objects for each object
patients <-
  list("pt18" = list("oHSV_d0" = objs[["pt18_pre"]],
                     "oHSV_d122" = objs[["pt18_post"]]),
       "pt22" = list("oHSV_d0" = objs[["pt22_pre"]],
                     "oHSV_d51" = objs[["pt22_post1"]],
                     "oHSV_d119" = objs[["pt22_post2"]]))

for (pt_name in names(patients)){
  # pull out one patient
  pt <-
    patients[[pt_name]]
  # normalize data, find variable features, scale data, and regress out mitochondrial genes
  for (sample in names(pt)){
    DefaultAssay(pt[[sample]]) <- "Spatial"
    pt[[sample]] <-
      pt[[sample]] %>%
      NormalizeData() %>%
      FindVariableFeatures() %>%
      ScaleData(vars.to.regress = "percent.mt")
  }
  # pt <-
  #   lapply(X = pt,
  #          FUN = SCTransform,
  #          vars.to.regress = "percent.mt",
  #          assay = "Spatial",
  #          return.only.var.genes = FALSE)
  # merge the objects for the patient
  spatial <-
    merge(x = pt[[1]],
          y = pt[2:length(pt)],
          add.cell.ids = names(pt))
  # get variable features for each sample for the patient and set them for the merged object
  var_feats <-
    SelectIntegrationFeatures(object = pt,
                              nfeatures = 4000)
  VariableFeatures(object = spatial) <- var_feats
  # normalize and scale spatial slot
  spatial <-
    NormalizeData(object = spatial,
                  assay = "Spatial") %>%
    ScaleData(assay = "Spatial",
              features = Features(spatial, 
                                  assay = "Spatial"))
  # normalize and scale protein slot
  spatial <-
    NormalizeData(object = spatial,
                  assay = "Protein",
                  normalization.method = "CLR") %>%
    ScaleData(assay = "Protein",
              features = Features(spatial, 
                                  assay = "Protein"))
  # perform dimensionality reduction and clustering on merged object
  DefaultAssay(object = spatial) <- "Spatial"
  spatial <- 
    spatial %>%
    RunPCA(assay = "Spatial",
           npcs = 30) %>%
    RunUMAP(dims = 1:30) %>%
    FindNeighbors(k.parm = 30)
  # make a umap to look at integration
  DimPlot(object = spatial,
          group.by = "orig.ident",
          cols = c("red","blue","orange"))
  ggsave(filename = paste0(fig, pt_name, "_merged_umap.jpeg"),
         width = 6,
         height = 5)
  # save object
  saveRDS(object = spatial,
          file = paste0(output, "rdata/", pt_name, "_pre_and_post_merged_log.rds"))
  # perform batch correction with harmony
  harmony <- 
    IntegrateLayers(object = spatial, 
                    method = HarmonyIntegration, 
                    normalization.method = "LogNormalize", 
                    verbose = F) %>%
    FindNeighbors(reduction = "harmony", 
                  dims = 1:30) %>%
    RunUMAP(dims = 1:30,
            reduction = "harmony")
  # make a umap to look at integration
  DimPlot(object = harmony,
          group.by = "orig.ident",
          cols = c("red","blue","orange"))
  ggsave(filename = paste0(fig, pt_name, "_harmony_umap.jpeg"),
         width = 6,
         height = 5)
  # join layers
  harmony[["Spatial"]] <- 
    JoinLayers(harmony[["Spatial"]])
  # save object
  saveRDS(object = harmony,
          file = paste0(output, "rdata/", pt_name, "_pre_and_post_merged_log_harmony.rds"))
}

# set final resolutions for each patient
harmony <-
  list("pt22" = readRDS(paste0(output, "rdata/pt22_pre_and_post_merged_log_harmony.rds")),
       "pt18" = readRDS(paste0(output, "rdata/pt18_pre_and_post_merged_log_harmony.rds")))
```